Packaging research artefacts with RO-Crate

### 2022-01-08 latex/ro-crate.tex
Sufficient (but not all) packages from ro-crate-proof

% add. options: [seceqn,secthm,crcready]
\documentclass[ds,crcready]{iosart2x}

@@

% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=darkerblue}

\usepackage{unicode-math}
\usepackage{graphicx}

@@

\DeclareGraphicsExtensions{.pdf,.png,.svg}
\graphicspath{../content/images/}
\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
\def\markdownOptionOutputDir{../output}
\def\markdownOptionHeaderAttributes{true}

% for table resizing
\usepackage{booktabs}
\usepackage{multirow}

@@

\pubyear{2022}
\volume{0}
\firstpage{1}
\lastpage{1}
\def\doi{10.3233/DS-210053}

@@

% Welcome to the 2000s with UNICODE fonts!
% Adapted from
% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
\setmathfont{STIX2Math}[
    Extension={.otf},
    Path=./STIX2fonts/,
    Scale=1
]
\setmainfont{STIX2Text}[
    Extension={.otf},
    Path=./STIX2fonts/,
    UprightFont={*-Regular},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic}
]

% https://typeof.net/Iosevka/ mono font has 
% all the logic symbols we need
%\setmonofont{iosevka-slab}[
%    Extension={.ttf},
%    Path=./ttf-iosevka-slab-6.1.3/,
%    Scale=MatchLowercase,
%    UprightFont={*-light},
%    ItalicFont={*-lightitalic}
%]

@@

%\begin{review}{editor}
%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
%\end{review}
%\begin{review}{solicited}
%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
%\end{review}

\begin{abstract}
\markdownInput{../content/01.abstract.md}
\end{abstract}

\begin{keyword}
\kwd{Data publishing}
\kwd{Data packaging}
\kwd{FAIR}
\kwd{Linked Data}
\kwd{Metadata}
\kwd{Reproducibility}
\kwd{Research Object}
\end{keyword}

\end{frontmatter}

@@

\textbf{Tool name} & \multicolumn{1}{c}{\tmultirow{2}{*}{Targets}}
& \tmultirow{2}{*}{Language/Platform} & \tmultirow{2}{*}{Status} \\

@@

Notable from this triple-like formalization is that a RO-Crate $R$ is
fully represented as a tree at depth 2 helped by the use of
%PDAG%
$\mathbh{IRI}$
%PDAG%
nodes. For instance the aggregation from the root entity
$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
File serialization is an equivalent shallow tree, although at depth 3
to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.

@@

%b22 ###bbsrt2
%b84 ###
\bibitem{doi:10.1007/978-3-030-45442-5_15}
%
\begin{bchapter}
\bauthor{\binits{J.}~\bsnm{Cardoso}},
\bauthor{\binits{D.}~\bsnm{Proen\c{c}a}} and
\bauthor{\binits{J.}~\bsnm{Borbinha}},
\bctitle{Machine-actionable data management plans: A knowledge retrieval
approach to automate the assessment of funders' requirements},
in: \bbtitle{Advances in Information Retrieval},
\beditor{\binits{J.M.}~\bsnm{Jose}},
\beditor{\binits{E.}~\bsnm{Yilmaz}},
\beditor{\binits{J.}~\bsnm{Magalh\~{a}es}},
\beditor{\binits{P.}~\bsnm{Castells}},
\beditor{\binits{N.}~\bsnm{Ferro}},
\beditor{\binits{M.J.}~\bsnm{Silva}} and
\beditor{\binits{F.}~\bsnm{Martins}}, eds,
\bpublisher{Springer International Publishing},
\blocation{Cham},
\byear{2020},
pp.~\bfpage{118}--\blpage{125}.
ISBN \bisbn{978-3-030-45442-5}.
\bid{doi={10.1007/978-3-030-45442-5_15}}
\end{bchapter}
%
\OrigBibText
J.~Cardoso,
D.~Proen{\c{c}}a and
J.~Borbinha,
\textit{Machine-Actionable Data Management Plans: A Knowledge Retrieval
Approach to Automate the Assessment of Funders' Requirements},
in: Advances in Information Retrieval,
J.M.~Jose,
E.~Yilmaz,
J.~Magalh{\~a}es,
P.~Castells,
N.~Ferro,
M.J.~Silva and
F.~Martins, eds,
Springer International Publishing,
Cham,
2020,
pp.~118--125.
ISBN 978-3-030-45442-5.
doi:10.1007/978-3-030-45442-5\_15.
\endOrigBibText
\bptok{structpyb}
\endbibitem

@@

%b104 ###bbsrt2
%b1 ###
\bibitem{sefton_blog_post_2021}
%
\begin{botherref}
\oauthor{\binits{P.}~\bsnm{Sefton}},
{FAIR} Data Management; It's a lifestyle not a lifecycle -- \surl{ptsefton.com},
2021.
\url{http://ptsefton.com/2021/04/07/rdmpic/}.
\end{botherref}
%
\OrigBibText
P.~Sefton,
{FAIR} Data Management; It's a lifestyle not a lifecycle - ptsefton.com,
2021.
\url{http://ptsefton.com/2021/04/07/rdmpic/}.
\endOrigBibText
\bptok{structpyb}
\endbibitem

>>>>

% add. options: [seceqn,secthm,crcready]
\documentclass[ds,v1.1.2,openaccess]{iosart2x}

@@

% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
%\usepackage{fontspec}
%\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=darkerblue}

%\usepackage{unicode-math}
%\usepackage{graphicx}

@@

\DeclareGraphicsExtensions{.pdf,.png,.svg}
\graphicspath{../content/images/}
%\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
%\def\markdownOptionOutputDir{../output}
%\def\markdownOptionHeaderAttributes{true}

% for table resizing
\usepackage{booktabs}
\usepackage{multirow}
%\usepackage{tmultirow}

@@

\volume{-1}
\issue{-1}
\pubyear{2022}
\firstpage{1}
\lastpage{42}
\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11

\startlocaldefs
\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
\endlocaldefs

@@

@@

\begin{review}{editor}
\reviewer{\fnms{Silvio} \snm{Peroni}\ead
[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
%}]{oe2}}
\end{review}

\begin{review}{solicited}
\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
\reviewer{\fnms{Alessia} \snm{Bardi}\ead
[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
\reviewer{\snm{one anonymous reviewer}}
\end{review}
%
%\dedicated{}
%
% HISTORY:
\received{\sday{13} \smonth{8} \syear{2021}}
\accepted{\sday{23} \smonth{9} \syear{2021}}
%
\begin{abstract}
%\parttitle{}
An increasing number of researchers support reproducibility by
including pointers to and
descriptions of datasets, software and methods in their publications.
However, scientific
articles may be ambiguous, incomplete and difficult to process by
automated systems. In
this paper we introduce RO-Crate, an open, community-driven, and
lightweight approach
to packaging research artefacts along with their metadata in a machine
readable manner.
RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
establish best
practices to formally describe metadata in an accessible and practical
way for their
use in a wide variety of situations.

An RO-Crate is a structured archive of all the items that contributed
to a research
outcome, including their identifiers, provenance, relations and
annotations. As a general
purpose packaging approach for data and their metadata, RO-Crate is
used across multiple
areas, including bioinformatics, digital humanities and regulatory
sciences. By applying
``just enough'' Linked Data standards, RO-Crate simplifies the process of
making research outputs FAIR while also enhancing research reproducibility.

An RO-Crate for this article\footnote{\url
{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
\end{abstract}
%
\begin{keyword} %% 1 u.c.
\kwd{Data publishing}
\kwd{data packaging}
\kwd{FAIR}
\kwd{Linked Data}
\kwd{metadata}
\kwd{reproducibility}
\kwd{research object}
\end{keyword}
%
\end{frontmatter}

@@

\textbf{Tool name} & \multicolumn{1}{c}{\multirow{2}{*}{Targets}}
& \multirow{2}{*}{Language/Platform} & \multirow{2}{*}{Status} \\

@@

Notable from this triple-like formalization is that a RO-Crate $R$ is
fully represented as a tree at depth 2 helped by the use of
$\mathbb{IRI}$
nodes. For instance the aggregation from the root entity
$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
File serialization is an equivalent shallow tree, although at depth 3
to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.

@@

%b22 ###bbsrt2
%b84 ###
\bibitem{doi:10.1007/978-3-030-45442-5_15}
%
\begin{bchapter}
\bauthor{\binits{J.}~\bsnm{Cardoso}},
\bauthor{\binits{D.}~\bsnm{Proen\c{c}a}} and
\bauthor{\binits{J.}~\bsnm{Borbinha}},
\bctitle{Machine-actionable data management plans: A knowledge retrieval
approach to automate the assessment of funders' requirements},
in: \bbtitle{Advances in Information Retrieval},
\beditor{\binits{J.M.}~\bsnm{Jose}},
\beditor{\binits{E.}~\bsnm{Yilmaz}},
\beditor{\binits{J.}~\bsnm{Magalh\~{a}es}},
\beditor{\binits{P.}~\bsnm{Castells}},
\beditor{\binits{N.}~\bsnm{Ferro}},
\beditor{\binits{M.J.}~\bsnm{Silva}} and
\beditor{\binits{F.}~\bsnm{Martins}}, eds,
\bpublisher{Springer International Publishing},
\blocation{Cham},
\byear{2020},
pp.~\bfpage{118}--\blpage{125}.
ISBN \bisbn{978-3-030-45442-5}.
\bid{doi={10.1007/978-3-030-45442-5\_15}}
\end{bchapter}
%
\OrigBibText
J.~Cardoso,
D.~Proen{\c{c}}a and
J.~Borbinha,
\textit{Machine-Actionable Data Management Plans: A Knowledge Retrieval
Approach to Automate the Assessment of Funders' Requirements},
in: Advances in Information Retrieval,
J.M.~Jose,
E.~Yilmaz,
J.~Magalh{\~a}es,
P.~Castells,
N.~Ferro,
M.J.~Silva and
F.~Martins, eds,
Springer International Publishing,
Cham,
2020,
pp.~118--125.
ISBN 978-3-030-45442-5.
doi:10.1007/978-3-030-45442-5\_15.
\endOrigBibText
\bptok{structpyb}
\endbibitem

@@

%b104 ###bbsrt2
%b1 ###
\bibitem{sefton_blog_post_2021}
%
\begin{botherref}
\oauthor{\binits{P.}~\bsnm{Sefton}},
{FAIR} Data Management; It's a lifestyle not a lifecycle -- ptsefton.com,
2021.
\url{http://ptsefton.com/2021/04/07/rdmpic/}.
\end{botherref}
%
\OrigBibText
P.~Sefton,
{FAIR} Data Management; It's a lifestyle not a lifecycle - ptsefton.com,
2021.
\url{http://ptsefton.com/2021/04/07/rdmpic/}.
\endOrigBibText
\bptok{structpyb}
\endbibitem
